Quantifying passive kelp rafting phenomena and connectivity of nearshore communities around the coast of South Africa through development of a 3D spectral numerical coastal-flow model

### 2020-06-29 thesis_chapter_1.Rmd
updates to outputs and rmd files for chapters 1,2 and 3. Updated figures to follow and final chpater 2 to follow

where $V_{curr}$ represents the current velocity relative to the earth, and $V_{rel}$ represents the object drift velocity relative to the ambient water. Ocean currents are determined by two components: the surface current (including the effects of Ekman drift, baroclinic motion, tidal and inertial currents) and Stokes drift induced by waves. The assumption made is $V_{curr}$ influences all floating objects in the same manner, regardless of shape or size. Therefore, $V_{curr}$ is equated with the surface current obtained from a numerical ocean model which has been parameterised by wind velocity. The effects of $V_{rel}$ on a floating object is driven by wind and wave forces which is dependent on the shape and size of the floating object. Previous research has shown that the effects of waves on drifting objects are negligible when the length of the object is less than the wave length; and increase significantly when the lengths are approximately the same [@grue1993; @hodgins1998]. Furthermore, current velocity $V_{curr}$ affects on drifting objects is considered to be negligible in this approach and ideally should be taken into account. Studies investigating kelp rafts have identified surface current velocity (as a function of wind velocity) as an important influence in determine trajectory [@harrold1989; @macaya2005; @hobday2000a; @hobday2000b]. Therefore, the objects motion/trajectory relative to the wind must be taken into account.

## Windage factors

An objects motion relative to the wind is sometimes referred to as "windage factors" or "leeway drift, and is difficult to accurately and empirically describe. The difficulty is due to accurate measurements needed for the current velocity, wind velocity, wave height and wave direction. Furthermore, these factors do not act independently and instead influence one another, which compounds the complexity of calculating wind effects on drifting objects. To overcome this challenge, measurements of current and wind velocity, and wave height and direction are measured along the objects trajectory for a wide range of conditions.

@@

Important to note is that the effects of wind and waves on the trajectory of a drifting object may be influenced by the object's orientation. This has been noted in modelling drift trajectories of ships [@allen1999review; @hackett2006], and should be incorporated into the leeway drift component. The direct ($V_l$) and indirect ($V_s$ and $V_w$) effects do not act independently and instead influence one another. with regards to this, @griffin2017 recommend incorporating an effective windage factor which would represent the combined effects of Stokes Drift, leeway drift and wave forces. Although a common approach, this may not be necessary in all circumstances particularly if individual effects of factors. As mentioned previously, the magnitude of effects of waves and wind on a drifting object is dependent on its shape, size and buoyancy.

@@

The buoyancy of objects at sea has been shown to be influeced by epibiont load [@hobday2000b; @rotha2011; @graiff2016; @macaya2016], which reduces drift times by reducing buoyancy with increasing biomass, which ultimatley leads to the object sinking. The bouyancy of the drifting object is determined by the growth rate of epiphytic species while drifting, which in turn will be depndent on environmental factors such as temperature and light which will vary along the objects trajectory Therefore, buoyancy should be parameterised by temporal epibiont biomass load.

>>>>

where $V_{curr}$ represents the current velocity relative to the earth, and $V_{rel}$ represents the object drift velocity relative to the ambient water [@hackett2006]. Ocean currents are determined by two components: the surface current (including the effects of Ekman drift, baroclinic motion, tidal and inertial currents) and Stokes drift induced by waves. The assumption made is $V_{curr}$ influences all floating objects in the same manner, regardless of shape or size. Therefore, $V_{curr}$ is equated with the surface current obtained from a numerical ocean model which has been parameterised on the wind velocity [@hackett2006]. The effects of $V_{rel}$ on a floating object is driven by wind and wave forces which is dependent on the shape and size of the floating object. In some cases the effects of wind or wave forces may be neglibible and only one aspect needs to be included or investigated seperately. Therefore, $V_{rel}$ can be seperated into wind and waves into two sub-components which allows the equation to be adpated to suite the particluar situation [@hackett2006]. For example, previous research has shown that the effects of waves on drifting objects are negligible when the length of the object is less than the wave length; and increase significantly when the lengths are approximately the same [@grue1993; @hodgins1998]. Furthermore, current velocity $V_{curr}$ affects on drifting objects is considered to be negligible in this approach and ideally should be taken into account. The literature identifies waves as the primary influencer of trajectory for larger objects while wind is regarded as the primary influencer for smaller objects. The effects of wind on object drift trajectory is complex and is ofcourse dependent on the exposed portion of the object to the wind. The effect of wind on trajectory is often referred to as leeway.

## Leeway

An objects motion relative to the wind is sometimes referred to as "windage factors" or "leeway drift", and is difficult to accurately and empirically describe. The difficulty is due to accurate measurements needed for the current velocity, wind velocity, wave height and wave direction. The appropriate approaches to measuring the afprmentioned parameters have been detailed by @allen1999 and is dependent on the objects size and shape as well as the quality and quantity of data available to the modeler. Additionally, these factors do not act independently and instead influence one another, which compounds the complexity of calculating wind effects on drifting objects. To overcome this challenge, measurements of current and wind velocity, and wave height and direction are measured along the objects trajectory for a wide range of conditions.

@@

Important to note is that the effects of wind and waves on the trajectory of a drifting object may be influenced by the object's orientation. This has been noted in modelling drift trajectories of ships [@allen1999review; @hackett2006], and should be incorporated into the leeway drift component. The direct ($V_l$) and indirect ($V_s$ and $V_w$) effects do not act independently and instead influence one another. with regards to this, @griffin2017 recommend incorporating an effective windage factor which would represent the combined effects of Stokes Drift, leeway drift and wave forces. Although a common approach, this may not be necessary in all circumstances particularly if individual effects of factors are being investigated. As mentioned previously, the magnitude of effects of wind and waves on an objects trajectory is depedent on the surface area/volume exposed to either influencer, which in turn is dependent on the bouyancy of the object. The object's buoyancy may also change over time due to other influencers such as erosion and epibiont biomass.

@@

The buoyancy is often a fixed value within a model and does not alter over time, such as @?? and @?? as examples. This may be because the effects of epibiont biomass on ships and wreckage is not regarded as a significant influencer. However, for smaller objects this may not be the case. This may be particulary true for living drifters such as macroalgae which are able to grow themselves, as well as provide a habitat for epiphytes and bryozoans which increase in biomass overtime. The buoyancy of objects at sea has been shown to be influeced by epibiont load [@hobday2000b; @rotha2011; @graiff2016; @macaya2016], which reduces drift times by reducing buoyancy with increasing biomass, which ultimatley leads to the object sinking. The bouyancy of the drifting object is determined by the growth rate of epiphytic species while drifting, which in turn will be depndent on environmental factors such as temperature and light which will vary along the objects trajectory Therefore, buoyancy should be parameterised by temporal epibiont biomass load.
